---
title: "R Notebook"
output:
  html_notebook: default
  word_document: default
---



```{r}
library(fpp3)
library(pdfetch)
library(pdfetch)
library(ggplot2)
library(tidyverse)
library(lubridate)
library(tsibble)
library(rugarch)
library(rmgarch)
library(fpp2)
library(forecast)
library(aTSA)
library(pdfetch)
library(fGarch)
library(stargazer)
library(dynlm)
library(urca)

```

```{r}
#get variables from world bank
inflation <- pdfetch::pdfetch_WB("FP.CPI.TOTL.ZG", c=("IND"))
#view(inflation)
unemployment <- pdfetch::pdfetch_WB("SL.UEM.TOTL.ZS", c=("IND"))
#view(unemployment)

```

```{r}
#rename variables

colnames(inflation) <- c("Inflation")
colnames(unemployment) <- c("Unemployment")

#INFLATION

#plot variables

autoplot(inflation) +
  ylab("Inflation %") +
  xlab("") +
  ggtitle("Inflation in India (CPI)")

```

Model: $\Delta  y_t =\gamma y_{t-1}+e_t$


```{r}
#test for unit roots 

adf_Inflation <- ur.df(inflation)
print(summary(adf_Inflation))

```
tau1 is the null hypothesis for $\gamma = 0$
 I get "Value of test-statistic is -2.1846" and the "Critical values for test statistics are: tau1 -2.6 -1.95 -1.61
Given that the test statistic is more extreme than critical value at 5%, we reject the null hypothesis that the unit root exist.


```{r}
#check for adf with drift 
adf_drift_inflation <- ur.df(inflation, type="drift")
print(summary(adf_drift_inflation))

```

model: $\Delta y_t = a_0 + \gamma y_{t-1}+e_t$ 
$a_0$ refers to the drift term in the model.
tau2 :Null hypothesis $\gamma=0$. (unit root)
phi1: Combines H0: $\gamma= a_0 = 0$(unit root but no drift)
Rejecting this null implies that one OR two was NOT zero.



For tau2, since the test statistics is -5.8262 is more extreme than critical values at 1%, 5%, and 10%.We reject the null hypothesis that there exist a unit root.


For phi1, since the test statistics is 16.9837 is greater than critical values at 1%, 5%, and 10%.We reject the null hypothesis that there exist a unit root and drift.


#check for adf with trend 
```{r}
adf_trend_inflation <- ur.df(inflation, type="trend")
print(summary(adf_trend_inflation))
```

Model: $\Delta y_t = a_0+\gamma *y_{t-1}+a_2*t+e_t$
$a_0$ refers to the drift term in the model.
tau3: Null hypothesis $\gamma=0$. (unit root)
phi3: Combined H0: $\gamma= a_0 = 0$(unit root but no drift)
phi2: H0: $\gamma= a_0 = a_2 = 0$(unit root, no drift term AND no time trend term)

For tau3, since the test statistics is -5.8218 is more extreme than critical values at 1%, 5%, and 10%.We reject the null hypothesis that there exist a unit root.

For phi2, since the test statistics is 11.3609 is more extreme than critical values at 1%, 5%, and 10%.We reject the null hypothesis that either gamma or a0 or both of them are not equal to zero.

For phi3, since the test statistics is 17.0303 is greater than critical values at 1%, 5%, and 10%.We reject the null hypothesis that either gamma or a0 or a2 or all of them are not equal to zero.

All three adf test shows that ou variable is stationary.

```{r}
#pp test for unit root 
pp_inflation <- pp.test(inflation)
pp_inflation
```


Null hypothesis: Data is not stationary
Since value of test statistics is -35.6812 which is more extreme than 4.4673, we can reject the null hypothesis and data is stationary.


pp test for trend 
```{r}
pp_inflation_trend <- ur.pp(inflation, model="trend")
print(summary(pp_inflation_trend))
```







Null hypothesis: Data is not stationary
Since value of test statistics is -35.3542 which is more extreme than 4.4673, we can reject the null hypothesis and data is stationary.
Since our both tests, ADF and PP are presenting the same results we don't need to implement the KPSS.


```{r}
#fit model - ARIMA - forecast 
fit_inflation <- auto.arima(inflation)
fit_inflation
library(forecast)
a=autoplot(forecast::forecast(fit_inflation, h=5))
as = forecast::forecast(fit_inflation, h=5)
plot(as$residuals)
qqnorm(as$residuals)
summary(fit_inflation)
```


```{r}
as = diff(unemployment,differences=4)
as = na.omit(as)
adf_unempolyment <- ur.df(as, type="none")
print(summary(adf_unempolyment))
```



```{r}

adf_unempolyment <- ur.df(unemployment, type="drift")
print(summary(adf_unempolyment))
```

```{r}

adf_unempolyment <- ur.df(unemployment, type="trend")
print(summary(adf_unempolyment))
```
```{r}
#pp test for unit root 
pp_unemployment <- ur.pp(unemployment,type = "Z-tau",model = "constant")
print(summary(pp_unemployment))
```

```{r}
#pp test for unit root 
pp_unemployment <- ur.pp(inflation,type = "Z-tau",model = "constant")
print(summary(pp_unemployment))
```

```{r}
autoplot(unemployment) +
  ylab("Inflation %") +
  xlab("") +
  ggtitle("Inflation in India (CPI)")
```


# Arima:

ACF plot is suggesting the value of q = 2 and value of p = 1
ARIMA should b 102

```{r}
acf(inflation)
pacf(inflation)
```


```{r}
fit_inflation <- auto.arima(inflation,ic='aic',trace = T)
fit_inflation

```
```{r}
fit_inflation <- arima(inflation, order = c(4,0,5))
```

```{r}
arch.test(fit_inflation)
```


```{r}
acf(ts(fit_inflation$residuals))
pacf(ts(fit_inflation$residuals))
```



```{r}
a = forecast::forecast(fit_inflation,level=c(95),h=5*1)
a
```

```{r}
autoplot(a)

```

















```{r}
sd<-diff(unemployment)
acf(sd,na.action = na.pass)
pacf(sd,na.action = na.pass)
```


```{r}
fit_unemployment<- auto.arima(sd,ic='aic',trace = T)
fit_unemployment
```



```{r}
acf(ts(fit_unemployment$residuals))
pacf(ts(fit_unemployment$residuals))
```



```{r}
a = forecast::forecast(fit_unemployment,level=c(95),h=5*1)
a
```

```{r}
autoplot(a)

```


























```{r}
Box.test(a$residuals,type = "Ljung-Box")
```

```{r}

```


```{r}
unemployment_model <- arima(unemployment, order = c(0,0,1), seasonal=list(order=c(0,0,0)))
summary(unemployment_model)
arch.test(unemployment_model)
tsdiag(unemployment_model)
```

```{r}
inflation_model <- arima(inflation, order = c(0,0,0), seasonal=list(order=c(0,0,0)))
arch.test(inflation_model)
summary(inflation_model)
tsdiag(inflation_model)
checkresiduals(inflation_model)
autoplot(inflation_model)
```



















```{r}
acf(inflation)
pacf(inflation)
```


```{r}
a <- arima(inflation, order = c(4,0,5))
```

```{r}
acf(ts(a$residuals))
pacf(ts(a$residuals))
```



# Arima:

```{r}
ad =  diff(unemployment,differences=4)
ad = na.omit(ad)

fit_unemployment <- auto.arima(ad,ic='aic',trace = T)
fit_unemployment
#stargazer(fit_inflation)
```
```{r}
acf(ts(fit_unemployment$residuals))
pacf(ts(fit_unemployment$residuals))
```



```{r}
a = forecast::forecast(fit_unemployment,level=c(95),h=5*1)
a
```

```{r}
autoplot(a)

```

```{r}
Box.test(a$residuals,type = "Ljung-Box")
```


Performs Portmanteau Q and Lagrange Multiplier tests for the null hypothesis that the residuals of a ARIMA model are homoscedastic.
```{r}
unemployment_model <- arima(unemployment, order = c(0,0,0), seasonal=list(order=c(0,0,0)))
arch.test(unemployment_model)
```




```{r}
library(ARDL)
dtmain = merge(unemployment, inflation, all = T)
dtmain = na.omit(dtmain)
dtmain =data.frame(dtmain)
library(strucchange)
sctest(Inflation~Unemployment,data =dtmain)

# ARDL model
mdl = auto_ardl(Inflation~Unemployment,data =dtmain,max_order = 5)
mdl
library(modelsummary)
modelsummary(mdl$best_model)
```


```{r}
library(lmtest)
dwtest(mdl$best_model)
```
```{r}
library(lmtest)
bptest(mdl$best_model)
```



```{r}
mdl = lm(Inflation~Unemployment,data=dtmain)
summary(mdl)
```

```{r}
grangertest(dtmain$Inflation~dtmain$Unemployment, order = 1)
```

